---
title: "OmicsBraid: Cross-Omics Effect Inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OmicsBraid: Cross-Omics Effect Inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(OmicsBraid)
```

## Why OmicsBraid?

Multi-omics studies frequently summarize each layer independently and then describe
RNA/protein/metabolite agreement qualitatively. OmicsBraid makes that comparison
explicit by estimating standardized effects with uncertainty, accounting for
matched-subject dependence, testing joint multi-omic evidence, quantifying
cross-layer heterogeneity, testing practical equivalence, and evaluating ordered
effect trajectories.

Version 0.2.2 retains the hierarchical inference introduced in v0.1.9, deliberately distinguishes **absence of detected evidence** from
**evidence of practical equivalence**, and separates confirmed braid patterns
from suggestive effect geometry.

## Simulate known biology

```{r simulate}
sim <- simulate_braid_data(n_per_group = 60, rho = 0.5, seed = 42)
sim$data
```

## Run the full workflow

For a quick vignette we use fewer bootstrap and Monte Carlo replicates than are
recommended for final analyses. Both the equivalence margin and the trajectory
margin are scientific SESOIs and should be justified in a real study.

```{r fit}
fit <- run_omics_braid(
  sim$data,
  group = "group",
  reference = "Control",
  comparison = "Disease",
  omic_order = c("RNA", "Protein", "Metabolite"),
  bootstrap_B = 150,
  ci_method = "percentile",
  integrated_ci_method = "percentile",
  ci_min_boot = 75,
  pattern_draws = 500,
  equivalence_margin = 0.30,
  trajectory_margin = 0.15,
  seed = 42
)
fit
```

## Layer-specific and integrated inference

```{r tables}
fit$effects
fit$integrated
fit$equivalence
fit$trend
fit$classification
```

The integrated table separates the multivariate all-zero omnibus test from the
GLS consensus effect. This is important when opposite layer effects cancel in the
consensus.


## Robust interval reporting

Version 0.2.2 can display matched subject-bootstrap intervals without changing the analytic SE/p-value inferential machinery. `fit$effects` and `fit$integrated` retain analytic interval columns alongside the selected percentile bootstrap intervals, while `fit$effect_intervals` and `fit$consensus_intervals` contain bootstrap diagnostics. BCa layer intervals are available for targeted/prefiltered analyses through `bootstrap_effect_intervals(..., method = "bca")`.

```{r robust-ci}
head(fit$effect_intervals)
head(fit$consensus_intervals)
```

## Ordered GLS trend test

`fit$trend` reports a covariance-aware slope over the prespecified omic order.
For same-direction layer effects, OmicsBraid only calls attenuation or
amplification when the slope exceeds the user-specified practical trajectory
margin with inferential support. Otherwise the broader concordant direction can
be retained with `pattern_status = "direction_confirmed"`.

## Null-like outcomes

`null_equivalent` means all observed layers passed practical-equivalence testing.
`no_detectable_effect` means the joint-null omnibus test was not rejected, but
the data were not precise enough to establish equivalence. These are different
scientific conclusions.

## Evidence Forest

```{r forest, fig.width=7, fig.height=4.5}
plot_evidence_forest(fit, "inversion")
```

## Effect Braid

```{r braid, fig.width=7, fig.height=4.5}
plot_effect_braid(fit, "inversion")
```

## Interpretation

A significant `Q_omics` indicates that the effect vector is poorly summarized by
a single common effect after accounting for estimated sampling covariance. The
I2-like index is descriptive and has no universal low/moderate/high cutoffs in
this development version.

Braid order is descriptive/inferential, not causal. A pattern such as `inversion`
describes supported effects across a prespecified layer order; it does not prove
molecular propagation. Likewise, `suggestive_pattern` describes effect geometry
when confirmatory precision is insufficient and must not be reported as a
confirmed result.
